\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
%\usepackage{tikz}
%\usetikzlibrary{shapes, shadows, arrows}
%\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
%\usepackage{pdflscape}
%\usepackage[flushleft]{threeparttable}
%\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
\begin{document}
\title{Regime Change}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk}} 
\date{\today}
\maketitle
\begin{abstract}
Hyman Minsky suggested that financial instability would proceed through an evolution from stability where lending was characterised by \emph{hedge financing} into more fragile regimes of \emph{speculative financing} and \emph{Ponzi financing}.    Though this process is not directly observable, there are outcomes that are more likely to be observed in each of these regimes.  Analysedk-down of \emph{uncovered interest partiy} (UIP) called the \emph{carry-trade} it is possible to identify these regimes.  Return characteristics change as financial instability develops so that the process can be modeled as a Markov chain where the states are unobserved but the outcome that is conditional on the state is used to uncover the parameters.  This is a version of the latent variable approach where the the states are related throuogh the Markov process. Identifying the financial state allows is used as a means of understanding more about financial instability. 

\end{abstract}

\section{The Minsky Model}
The Minsky model of endogenously increasing financial fragility has rightly received increased attention in the aftermath of the 2007 - 2008 global financial crisis.  The model suggest that a period of economic stability will provide the condidtions for a gradual transition of lending within the economy from one characterised as \emph{hedge} though that which is more \emph{speculative} and into something that can be desribed as {Ponzi}.  Under hedge financing the level of debt will not tend to extend beyond the ability to repay interest and principal; under specuative lending debt-to-equity ratios start to rise and the level of borrowing relative to income is elevated to a point where it becomes difficult to repay principal without asset price apprecation; under the Ponzi lending there is insufficient revenue to repay principal or interest and asset price appreciation or increased borrowing is necessary to mantain the debt. 

While this is a plausible description of rising financial fragility, it is often difficult to determine at what position the economy is in the financial cycle.  While measuring debt-to-equity ratios and the scale of bank lending may provide some indication about the regime that is in place, it is clear that financial services business evolve in ways that make loand counting in inadequate. The recent financial crisis showed that innovations like \emph{Collateralised Debt Obligations (CDO)} can allow an increase in debt that does not directly in bank lending.  

\section{Hidden Markov Models}
These models are used extensively in biometrics to analyse genome structure.  In particular, they are used to assess the underlying regime that is influencing the gene strcture. 

\href{http://a-little-book-of-r-for-bioinformatics.readthedocs.org/en/latest/src/chapter10.html}{Chapter 10 Biometric Text on HMM}
has an excellent overview of markov HMM and the R code necessary. One component of this that could be of interest is the assertion in on-line bimetrixs text that it is a problem to find the underlying state that produced the DNA outcome.  The equivalent of this for the crash model is to find the underlying Minsky state that produced the market activity. 

This is the folder for builing the regime-switching model. The three latent states are the periods of shock, of calm and of crisis.  These are unseen but may be identified by other variables (such as the level of international risk aversion (VIX) or the state of domestic political uncertainty (see that buy at Yale???)).  Alternatively, it may be possible to draw the states from the data and compare the information that is supplied by the data with that from what is known about politicla and economic developments at the time.  It is also possible to assess the probability that there will be a switch from one regime to another.  This is a Markov-swithing model. 

The motivation for this research is to help to identify regimes and therefore help authorities determine the risk of financial crisis or crash. Use the FX market to get a broader understanding of some of the forces at work.

From Maatin's speech (can be deleted later).  In a \emph{mixture model}, each observation is assumed to be drawn from a number of distinct sub populations.  These can be called \emph{component distributions}.  The distribution from which the component is drawn is not immediately observable and is therefore represented as a \emph{latent state}. 

A mixture distribution is defined as 
\begin{equation}
p(Y_1 = y) = \sum_{i - 1}^N p(Y_t = y|S_t = i)P(S_t = i)
\end{equation}
where,
\begin{itemize}
\item $S_t \in {1, \dots, N}$ denotes the latent state or class of observation t
\item $P(S_t = i)$ denots the probability of the latent state t equals i 
\item $p(Y_t = y|S_t = i)$ denotes the density of observation of $Y_t$ conditional on latent state being $S_t = i$.
\end{itemize}
\section{Perth Water Example}

In the Perth water example that Maatin presented, there are three different models (level, linear and quadratic).  The change model assumes that there is one or more discrete change points.  It may be the mean, trend or other parameters that change. The transition matrix defines the change points.  

\begin{equation*}
\begin{pmatrix}
p_1 & 1 - p_1 \\
0 & 1
\end{pmatrix}
\end{equation*}
Where $p_1$ is the probability that the system will be in state 1.  One there is a switch to state two.  This matrix can be extended for multiple states. What is interesting in the modesl or Perth water flow are that there is a transition to a second state where there is a lower mean and a lower standard deviation. Therefore, the water level is lower and the variability of the water is less. 

Ask Maarten how it determines whether the states can switch back and forth or stay fixed?  It looks as if the answer to this question is the transition matrix can be restricted to ensure that this happens.  On page 21 of Maarten's report, it says that, in the EM algorithum, probabilities initialised as 0 or 1 will remain there.  However, this does not happen with the bull and bear market example.  Why???

There is an example of this method in the analysis of SP 500 regimes: Gains with low volatuility and losses with high volatility. The model will identify the most likely 

It may be possible to start from the distribution of the returns in the period of crisis and the period of calm and then to move to the identification of transition process.  This would be like the slide 9 of Maarten's London R presentation. 

In the \emph{dependent mixture model} states are assumed to be statistically dependent.  This is consistent with the Minsky theory that the period of calm creates the conditions for the crash. The process underlying the state transitions is a \emph{homogenous first order Markov process}  (look this up for additional definition).  This process is completely defined by the initial state probabilities.  

\begin{equation*}
P(S_1 = 1), \dots P(S_1 = N)
\end{equation*}
and the state transition matrix, 

\begin{equation*}
\begin{pmatrix}
P(S_t = 1|S_{t-1}=1) & P(S_t = 2|S_{t-1}=1) & \dots & P(S_t = N|S_{t-1}=1)\\
P(S_t = 1|S_{t-1}=2) & P(S_t = 2|S_{t-1}=2) & \dots & P(S_t = N|S_{t-1}=2)\\
\vdots & \vdots & \ddots & \vdots \\
P(S_t = 1|S_{t-1}=N) & P(S_t = 2|S_{t-1}=N) & \dots & P(S_t = N|S_{t-1}=N)
\end{pmatrix}
\end{equation*}

The models are estimated using the Expectation-Maximisation (EM) or numerical optimisation (when parmeters are constrained).  The dependent mixture model is made up of three sub models:  
\begin{enumerate}
\item The prior model: $P(S_1|x, \theta_{prior})$
\item The transition model: $P(S_t|x, S_{t-1}, \theta_{trans})$
\item The response model: $P(Y_t| S_t, x, \theta_{resp})$
\end{enumerate}
 See Maarten's notes to see how these are implemented.  

\section{Literature Review}
The use of the VIX index to identify times of risk aversion can make the period of crisis an observable state of the world. This is about a regime-switching model. This will treat the states of the world to be unobservable and to fit a regime-switching model (build like that of Hamilton (1989).  These notes from Ghysels.  There are two states of the world:  crisis and moderation.  If the system is in a crisis, it stays there with probability p; it switches to moderation with probability $1-p$.  If in moderation, the system stays there with a probability q and switches to crisis with probability $1-q$.  If the probabilities change over time, there is no longer a \emph{homogenous Markov Chain}. Ghysels has a seasonal dummy for the probabilities that represent the months or quarters.  

\href{http://members.home.nl/jeroenvermunt/dias2010.pdf}{Mixture Hidden Markov Models} Hidden models help to calafify the regime under which securities trade. Model takes into account the unobserved hetrogeneity across time. This could could be extended to space (for different countries).  

\section{Methods}
\subsection{Data}
The data are a sample of CEE carry trades that have been compiled from raw exchange rate and interest rate data.  They show a range of possible carry trades that could have been conducted. 

The data are constructed with the function \emph{forp} in the \emph{Raw.R} file. This programme will create a sample of carry-trade profits from the inputs that will provide a funding currency from the set of US dollar, Euro, Swiss Franc and Japanese yen; an investment currency from the set of CEE countries provided; a time period of 1 month or 3 months.  Is there any need for any others funding curencies?  I don't think so!   It is possible to add other investment currencies.  There is ISK, TRY and NOK as reference.  It would be possible to have shorter time periods for the carry trade.   This would require the addition of the appropriate times series for LIBOR or deposit rates.  

The forumla for the carry can be taken from the doctorate. 

\subsection{Regime selection}
There are three states or regimes: the first is caution, the second is build and the third is caution. These three regimes or states loosely correspond to the three stages of financial instability: hedge finance, speculative finance and Ponzie finance. It is assumed that the Ponzie stage swiftly turns into a crash. 

Maybe put the R details in an appendix. 

The returns are modeled as a simple model of the level of returns and the standard deviation around this level.  it is assumed that there are three different regimes so the level and the standard deviation around this level is allowed to change with the regime.  

Three regimes are identfied with the 1-month investment in Polish zloty funded by US dollars.  These regimes are:  caution, carry-build and crash. 

\subsection{Wikipedia}
From wikipedia.  \href{http://en.wikipedia.org/wiki/Hidden_Markov_model}{Hidden Markov Models}.  The system is a Markov process.  There are unobserved, hidden states.  The Markov property A Markov chain satisfied the Markov property.  That means that it depends only on the current state.  It has no memory. \emph{Brownian motion}, for example, is a Markov process. In a Markov chain, the state is observed and the transition probabilities are the onl;y parameter.  In a HMM, each state has a probability distribution over the possible outcomes.  These can be called \emph{tokens}.  A series of \emph{tokens} can tell us someting about the states.   A hidden Markov model is a generalisation of the mixture model where hidden or latent variables are related through a Markov process. These latent variables comntrol the outcome token. A mixture model is a model with sub-populations where the observations do not allow identification of the sub populations.  A mixture model has a mixture distribution. A mixture distribution is a distribtion of a random variable that is derived from a number of random variales that are related to each other. If the variable is continuous, this is a mixture density.  The \emph{mixture components} are combined together to form the mixture distribution with certain \emph{mixture weights}.  Discrete are sometimes call \emph{compound distributions}.  There is a difference between adding together two normal distributons (where there will be a new normal with a new mean) and the mixture distibution (where there will be twin peaks).  

\href{http://members.home.nl/jeroenvermunt/dias2010.pdf}{Mixture Hidden Markov Models}  The observed response $y_{it}$ is the return of stock market $i$ at time $t$.  There are also two latent variables:  a time-constant discrete latent variable and a time-varying discrete latent variable.  The former is denoted by $w \in {1, \dots, S}$ captures unobserved hetoegeneity across markets; $z_t \in {1,2}$ is a two-state, time-varying latent variable. 

$f(y_{it}|z_t)$ is assumed to have a multivariate normal density function. This distribution is characterised by $\theta_k = (\mu_k, \sigma_k^2)$.  Excluding the states $w$, there are the initial state probabilities to be determined, the 2 transition probabilities and the conditional mean and conditional variance to be estimated.  Thsi is done by Maximum likelihood using the log-likelihood function $l(\varphi, y) = \sum_{i=1}^n log f(y_i; \varphi)$. This is a problem that can be solved with the \emph{Expectation-Maximization (EM) algorithm}.  Dempster, Laird and Rubin (1977).  The E step computes the joint conditional distribution of the latent variables given the data and the current provisional estimates of the model parameters. The M step ML methods are used to update the parameters using the estimated densities of the latent variables as weights. For hidden Markov models, as special variant of the EM algorithm is proposed (called \emph{the forward-backward} or \emph{Baum-Welch} alogorithm (Baum et al 1970).   

The markets are categorised as either \emph{bull} or \emph{bear} markets. This is consistent with the return volatility of markets.  About 25\% of the returns are in the bear markets. The evidence that the culster of markets that are closer to emerging markets are more liekly to be in a bear market than the devloped markets. In addition, it shows that there is persistence.  Once a market is in a particular regime (z) it is likely to stay there. Cluster 2 shows a lower propensity to move to a bear market from a bull market.  This is consistent with what is known about emerging markets.  The picture of the estimated posterior bull market regime is shown for emerging and deveped markets.  It shows that emerging markets are more likely to suffer bear markets.  




\section{Results}
Three regimes can be identified: number one, the cautious period; number two, the time when carry positions are being built; number three, the crash.  This works for PLNUSD, HUFEUR. 
\end{document}